Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-30 PAPER.tex
made all of Jonnie's minor edits

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, error estimation, and bias) that are essential to investigate 
for a robust 21\,cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned. A companion paper, Kolopanis et al. (\textit{submitted}), builds off of the methods in this paper to present PAPER-64 results for multiple redshifts and baseline types.

Finally, this paper adds to the growing foundations of lessons which have been documented in \citet{Paciga2013}, \citet{Patil2016}, and \citet{Jacobs2016}, by the GMRT, LOFAR, and MWA projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

@@

\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
Here, \noindent $\textbf{Q}$ is a family of matrices that takes our frequency-domain visibilities and Fourier transforms them, while also converting from Jy to 
Kelvin and taking into account cosmological scalings. It is formally evaluated as $\textbf{Q}^{\alpha} \equiv \frac{\partial\textbf{C}}{\partial p_{\alpha}}$ (which has dimensions ($N_{f}$,$N_{f}$)), or the derivative of the covariance $\textbf{C} \equiv \langle\vec{\textbf{x}}\vec{\textbf{x}}^{\dagger}\rangle$ with respect to the true bandpower $p_\alpha$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the 
Fourier dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}), which is a good approximations for the short baselines that PAPER analyzes). $\textbf{R}$ is a weighting matrix with dimensions ($N_{f}$,$N_{f}$) --- as 
an example, inverse covariance weighting (the optimal form of QE) would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $
\textbf{R} \equiv \textbf{I}$, the identity matrix.

@@

\begin{equation}
\label{eq:phat}
\widehat{\textbf{p}} = \textbf{M}\widehat{\textbf{q}},
\end{equation}

\noindent where $\widehat{\textbf{p}}$ is the estimate of the true power spectrum $\textbf{p}$. The data analyst has a choice for $
\textbf{M}$. For simplicity in this section we choose $\textbf{M}$ to be diagonal, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}, subject to the constraint that each element of $\widehat{\textbf{p}}$ is a weighted sum of the elements of $\textbf{p}$ with weights that sum to unity.

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectrum estimates $\widehat{\textbf{p}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift-scan observation. Hence, flat spectrum sources (away 
from zenith) oscillate in time and frequency in our measurements. We therefore form a mock bright foreground signal, $
\textbf{x}_{\rm FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock EoR signal, $\textbf{x}_{\rm EoR}$, as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $p(k)$ (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, here we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data matrix is then $\textbf{x} = 
\textbf{x}_{\rm FG} + \textbf{x}_{\rm EoR}$, to which we apply different weighting schemes throughout Section \ref{sec:SiglossOverview}. The three data components are shown in Figure 
\ref{fig:toy_sigloss1}.

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss1.pdf}
	\caption{Our toy model dataset to which we apply different weighting schemes to in order to investigate signal loss. We model a mock foreground-only visibility with a sinusoid signal that varies smoothly in 
time and frequency. We model a mock EoR signal as a random Gaussian signal. We add the two together to form $\textbf{x} = 
\textbf{x}_{\rm FG} + \textbf{x}_{\rm EoR}$. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

Altering $\widehat{\textbf{C}}$ as such is one specific example of a regularization method for this toy model, in which we are 
changing $\widehat{\textbf{C}}$ in a way that changes its coupling to the data realization. The resulting 
power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover the EoR signal, 
demonstrating that if we can disentangle the foreground-dominating modes from EoR-dominating modes, we can down-weight 
them with negligible signal loss. There are several other ways to regularize $\widehat{\textbf{C}}$, and we will discuss some in Section 
\ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to the coupling of weak eigenmodes (EoR-dominated modes) to the data. We will next show how this effect is exaggerated by reducing the total number of independent 
samples in a dataset.

@@

Although smooth-spectrum foregrounds preferentially show up at low delay, or low $k$-modes, their isolation within the wedge is not perfect. In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the delay associated with the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum, the signal chain, and the 
antenna response, all of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017}.

@@

\begin{equation}
\Delta^{\textbf{2}}(k) = \frac{k^{3}}{2\pi^{2}}\,\hat{p}(k),
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{\rm out,\alpha} &\equiv& \widehat{\textbf{p}}_{r,\alpha} \nonumber\\%-\widehat{\textbf{p}}_{x,\alpha} \nonumber \\
&=& \text{M}^{\alpha}_{r}\textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r},% - \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x},
\end{eqnarray}
where, for illustrative purposes and notational simplicity, we have written these equations with scalar normalizations M, even though for our numerical results we choose a non-diagonal matrix normalization using $\mathbf{M}$ as in Equation \eqref{eq:phat}.

@@

%\begin{equation}
%P_{\rm out} = \widehat{\textbf{p}}_{x} \label{eqn:anzatz},
%\end{equation}
%and folding in our definition from Equation \eqref{eq:sigloss}, this ansatz implies that we are seeking the injected 
%signal amplitude which causes a power doubling ($\phat_r = 2 \phat_x$). \cc{I am unsure of this power doubling thing}

@@

%In short, we can approximate a signal loss estimate as the ratio of $P_{\rm out}/P_{\rm in}$, evaluated at the data level $\widehat{\textbf{p}}_{x}$. We motivate the fact that we can evaluate the output-to-input power spectrum ratio at $\widehat{\textbf{p}}_{x}$ by the following reasoning (and then detail our signal loss estimation in practice in the sections that follow).

@@

\begin{eqnarray}
\label{eq:crossterm_full}
P_{\rm out,\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) \nonumber \\%- 
%\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms can be summed together as:

@@

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are Gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniformly weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a symmetric distribution with zero mean and width $\sqrt{p_e}\sqrt{p_x}$. We also note that in this regime, $\widehat{\textbf{p}}_{r}$ (black) approaches the data level (gray) as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $P_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $P_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the estimated inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a symmetric distribution of zero mean and width $\sqrt{p_e}\sqrt{p_x}$.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.pdf}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. The net output ($P_{\rm out}=\widehat{\textbf{p}}_r$) is shown in black, the individual terms, as defined by Equation 
\eqref{eq:crossterm},  are shown in blue, red, green; the dotted black line indicates 1:1 input to output mapping.  The 
empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015} is on the left and the uniform-weighted 
case is on the right. The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.   The gray 
horizontal line is the power spectrum value of data alone (doesn't depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than the net output because the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant, the cross-terms have a noise-like 
term with width $\sqrt{p_e}\sqrt{p_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for 
the weighted case). }
	\label{fig:sigloss_terms}
\end{figure*}

@@

Going forward, we treat every $k$-value separately in order to determine an upper limit on the EoR signal per $k$. We bin our simulation outputs along the $P_{\rm in}$ axis, and smooth the distribution of $P_{\rm out}$ values using kernel density estimators (\citealt{scott2008kernel}) for each bin (and normalizing them). Stitching all of them together results in a 2-dimensional smoothed transfer function, shown in Figure \ref{fig:sigloss_transfercurve} for both the weighted (left) and unweighted (right) cases. Although we only show figures for one $k$-value, we note that 
the shape of the transfer curve is similar for all $k$'s. This transfer function is the likelihood function in Bayes' rule, namely $\mathcal{L} (  P_{\rm out} | P_{\rm in})$.  We then invoke Bayes' interpretation and re-interpret it as the posterior $p(P_{\rm in}|P_{\rm out})$ where we recall that $P_{\rm out}$ is a model of our data. To do this we make a horizontal cut across at the data value $\widehat{\textbf{p}}_{x}$ (setting $P_{\rm out} = \widehat{\textbf{p}}_{x}$) to yield a posterior distribution for the signal. We compute the $95\%$ confidence interval (an upper limit on EoR) from these distributions.

@@

By-eye inspection of the transfer function in Figure \ref{fig:sigloss_transfercurve} gives a sense of what the signal loss result should be. The power spectrum value of our data, $
\widehat{\textbf{p}}_{x}$ is marked by the solid gray horizontal lines. From the left plot (empirically estimated inverse covariance weighting), one can eyeball that a data value of $10^{4} \,$mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, for example, would map approximately to an upper limit of $\sim10^{8} \,$mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, implying a signal loss factor of $\sim10^{4}$. For the uniform-weighted case (right plot), we see no loss at a data value of $\sim10^{7} \,$mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$.

@@

\begin{equation}
\label{eq:sense}
n(k) = \frac{X^{2}Y \Omega_{\rm eff} T_{\rm sys}^{2}}{\sqrt{2N_{\rm lst}N_{\rm seps}}\,t_{\rm int}N_{\rm days}N_{\rm bls}N_{\rm pols}}.
\end{equation}
We will now explain each factor in Equation \eqref{eq:sense} and highlight key differences from the numbers used in \citetalias{ali_et_al2015}.

>>>>

The major challenge that faces all 21\,cm experiments is isolating a small signal that is buried underneath foregrounds and 
instrumental systematics that are, when combined, four to five orders of magnitude brighter \citep[e.g.,][]{santos_et_al2005, ali_et_al2008, deOliveiraCosta_et_al2008, jelic_et_al2008, bernardi_et_al2009, bernardi_et_al2010, ghosh_et_al2011, pober_et_al2013b, bernardi_et_al2013, dillon_et_al2014, kohn_et_al2016}. A clean measurement therefore requires an intimate understanding of the instrument and a rigorous study of data analysis choices. With continual progress being made 
in the field and HERA on the horizon, it is becoming increasingly important to understand how the methods we choose interact 
with each other to affect power spectrum results. More specifically, it is imperative to develop techniques and tests that ensure 
the accuracy and reliability of a potential EoR detection. In this paper, we discuss three topics (signal loss, error estimation, and bias) that are essential to investigate 
for a robust 21\,cm power spectrum analysis. We also highlight four power spectrum techniques (fringe-rate filtering, weighting, bootstrapping, jackknife testing) and their trade-offs, potential 
pitfalls, and connections to the themes. We first approach the themes from a broad perspective, and then perform a detailed 
case study using data from the 64-element configuration of PAPER, motivating a revised PAPER-64 power spectrum 
from the lessons learned. A companion paper, Kolopanis et al. (\textit{submitted}), builds off of the methods in this paper to present PAPER-64 results for multiple redshifts and baseline types.

Finally, this paper adds to the growing foundations of lessons which have been documented, for example, in \citet{Paciga2013}, \citet{Patil2016}, and \citet{Jacobs2016}, by the GMRT, LOFAR, and MWA projects respectively. These lessons are imperative as the community as a whole moves towards higher sensitivities and potential EoR detections.

@@

\begin{equation}
\label{eq:qhat}
\widehat{q}_{\alpha} = \frac{1}{2}\textbf{x}^{\dagger}\textbf{R}\textbf{Q}^{\alpha}\textbf{R}\textbf{x}.
\end{equation}
Here, \noindent $\textbf{Q}$ is a family of matrices that takes our frequency-domain visibilities and Fourier transforms them, while also converting from Jy to 
Kelvin and taking into account cosmological scalings. It is formally evaluated as $\textbf{Q}^{\alpha} \equiv \frac{\partial\textbf{C}}{\partial p_{\alpha}}$ (which has dimensions ($N_{f}$,$N_{f}$)), or the derivative of the covariance $\textbf{C} \equiv \langle\textbf{x}\textbf{x}^{\dagger}\rangle$ with respect to the true bandpower $p_\alpha$, where $\alpha$ indexes a waveband in $k_{\parallel}$ (a cosmological wavenumber $k_{\parallel}$ is the 
Fourier dual to frequency under the delay approximation (\citealt{parsons_et_al2012b}), which is a good approximation for the short baselines that PAPER analyzes). $\textbf{R}$ is a weighting matrix with dimensions ($N_{f}$,$N_{f}$) --- as 
an example, inverse covariance weighting (the optimal form of QE) would set $\textbf{R} \equiv \textbf{C}^{-1}$ and a uniform-weighted case would use $
\textbf{R} \equiv \textbf{I}$, the identity matrix.

@@

\begin{equation}
\label{eq:phat}
\widehat{\textbf{P}} = \textbf{M}\widehat{\textbf{q}},
\end{equation}

\noindent where $\widehat{\textbf{P}}$ is the estimate of the true power spectrum $\textbf{P}$. The data analyst has a choice for $
\textbf{M}$. For simplicity in this section we choose $\textbf{M}$ to be diagonal, although we explore other cases for the analysis 
of PAPER-64 data as explained in Section \ref{sec:Bias}, subject to the constraint that each element of $\widehat{\textbf{P}}$ is a weighted sum of the elements of $\textbf{P}$ with weights that sum to unity.

In the next three sections, we use toy models to investigate the effects of weighting matrices on signal loss by experimenting with different matrices $\textbf{R}$ and examining their impact on the resulting power spectrum estimates $\widehat{\textbf{P}}$. Our goal in 
experimenting with weighting is to suppress foregrounds and investigate EoR losses associated with it. We note that we 
purposely take a thorough and pedagogical approach to describing the toy model examples given in the next few sections. The specifics of how signal loss appears in PAPER's analysis is later described in Section 
\ref{sec:Sigloss}.

@@

We create mock visibilities, $\textbf{x}$, and assume a non-tracking, drift-scan observation. Hence, flat spectrum sources (away 
from zenith) lead to measured visibilities which oscillate in time and frequency. We therefore form a mock visibility measurement of a bright foreground signal, $
\textbf{x}_{\rm FG}$, as a complex sinusoid that varies smoothly in time and frequency, a simplistic but realistic representation of a 
single bright source. We also create a mock visibility measurement of an EoR signal $\textbf{x}_{\rm EoR}$ as a complex, Gaussian random signal. A more 
realistic EoR signal would have a sloped power spectrum in $p(k)$ (instead of flat, as in the case of white noise), 
which could be simulated by introducing frequency correlations into the mock EoR signal. However, here we treat all 
$k$'s separately, so a simplistic white noise approximation can be used. Our combined data matrix is then $\textbf{x} = 
\textbf{x}_{\rm FG} + \textbf{x}_{\rm EoR}$, to which we apply different weighting schemes throughout Section \ref{sec:SiglossOverview}. The three data components are shown in Figure 
\ref{fig:toy_sigloss1}.

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip,width=\columnwidth]{plots/toy_sigloss1.pdf}
	\caption{Our toy model dataset to which we apply different weighting schemes to in order to investigate signal loss. We model a mock foreground-only visibility with a sinusoid signal that varies smoothly in 
time and frequency. We model a mock visibility of an EoR signal as a random Gaussian signal. We add the two together to form $\textbf{x} = 
\textbf{x}_{\rm FG} + \textbf{x}_{\rm EoR}$. Real parts are shown here.}
	\label{fig:toy_sigloss1}
\end{figure}

@@

Altering $\widehat{\textbf{C}}$ as such is one specific example of a regularization method for this toy model, in which we are 
changing $\widehat{\textbf{C}}$ in a way that changes its coupling to the data realization. The resulting 
power spectrum estimate for this case is shown in the right plot of Figure \ref{fig:toy_sigloss3}. In this case we recover the EoR signal, 
demonstrating that if we can disentangle the foreground-dominating modes and EoR-dominating modes, we can down-weight 
them with negligible signal loss. There are several other ways to regularize $\widehat{\textbf{C}}$, and we will discuss some in Section 
\ref{sec:otherweight}.

@@

We have shown how signal loss can arise due to the coupling of weak eigenmodes (EoR-dominated modes) to the data. We will next show how this effect is exacerbated by reducing the total number of independent 
samples in a dataset.

@@

Although smooth-spectrum foregrounds preferentially show up at low delay, or low $k$-modes, their isolation within the wedge is not perfect. In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the delay associated with the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum, the signal chain, and the 
antenna response, all of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017, kerrigan_et_al2018}.

@@

\begin{equation}
\Delta^{\textbf{2}}(k) = \frac{k^{3}}{2\pi^{2}}\,\hat{P}(k),
\end{equation}

@@

\begin{eqnarray}
\label{eq:sigloss}
P_{\rm out,\alpha} &\equiv& \widehat{\textbf{P}}_{r,\alpha} \nonumber\\%-\widehat{\textbf{P}}_{x,\alpha} \nonumber \\
&=& \text{M}^{\alpha}_{r}\textbf{r}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{r},% - \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x},
\end{eqnarray}
where, for illustrative purposes and notational simplicity, we have written these equations with scalar normalizations M, even though for our numerical results we choose a non-diagonal matrix normalization using $\mathbf{M}$ as in Equation \eqref{eq:phat}.

@@

%\begin{equation}
%P_{\rm out} = \widehat{\textbf{P}}_{x} \label{eqn:anzatz},
%\end{equation}
%and folding in our definition from Equation \eqref{eq:sigloss}, this ansatz implies that we are seeking the injected 
%signal amplitude which causes a power doubling ($\phat_r = 2 \phat_x$). \cc{I am unsure of this power doubling thing}

@@

%In short, we can approximate a signal loss estimate as the ratio of $P_{\rm out}/P_{\rm in}$, evaluated at the data level $\widehat{\textbf{P}}_{x}$. We motivate the fact that we can evaluate the output-to-input power spectrum ratio at $\widehat{\textbf{P}}_{x}$ by the following reasoning (and then detail our signal loss estimation in practice in the sections that follow).

@@

\begin{eqnarray}
\label{eq:crossterm_full}
P_{\rm out,\alpha} &=& \text{M}^{\alpha}_{r}(\textbf{x}+\textbf{e})^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}(\textbf{x}+\textbf{e}) \nonumber \\%- 
%\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x} \nonumber \\
&=& \text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} + \text{M}^{\alpha}_{b}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}
^{\alpha}\textbf{R}_{r}\textbf{e} \nonumber \\
&+& \text{M}^{\alpha}_{c}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{e} + \text{M}^{\alpha}_{d}\textbf{e}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} %\nonumber \\
%&-& \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}.
\end{eqnarray}
Assuming \textbf{R}$_{r}$ is symmetric, the two cross-terms (terms with one copy of $\textbf{e}$ and one copy of $\textbf{x}$) can be summed together as:

@@

{\bf Small injection:}
%In this regime, the first term in Equation \eqref{eq:crossterm} is very similar to the power spectrum of data alone, namely $\text{M}^{\alpha}_{a}\textbf{x}^{\dagger}\textbf{R}_{r}\textbf{Q}^{\alpha}\textbf{R}_{r}\textbf{x} \sim \text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{R}_{x}\textbf{Q}^{\alpha}\textbf{R}_{x}\textbf{x}$. They differ only in their covariance matrices, and when the injected signal is very small compared to the data, their difference can be neglected. 
%This is shown by the black and gray curves converging in both panels of Figure \ref{fig:sigloss_terms}. 
In this regime, the cross-terms (red) behave as noise averaged over a finite number of samples. Output values are Gaussian distributed around zero, spanning a range of values set by the injection level. This is because $\widehat{\textbf{R}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below). In fact, for the uniformly weighted case, the cross-term  $\text{M}^{\alpha}_{x}\textbf{x}^{\dagger}\textbf{I}\textbf{Q}^{\alpha}\textbf{I}\textbf{e}$ is well modeled as a symmetric distribution with zero mean and width $\sqrt{P_e}\sqrt{P_x}$. We also note that in this regime, $\widehat{\textbf{P}}_{r}$ (black) approaches the data level (gray) as expected. %In this limit we might imagine that the cross-correlation should also be small, enabling us to recover just the EoR term. However, examining the cross-term plotted in Figure \ref{fig:sigloss_terms} (red) in the identity case (right plot), we see that this term is still large enough to have a significant impact on $P_{\rm out}$. It is in this regime where we're dominated by biases from the cross-terms, causing the difference between the signal-only term (green) and $P_{\rm out}$ (black). We also note that the cross-terms behave as noise at small injections even in the inverse covariance weighted case (left plot), spanning a range of values set by the injection level. This is because $\widehat{\textbf{C}}_{r}$ is dominated by the data $\textbf{x}$, avoiding correlations with $\textbf{e}$ that can lead to solely negative power (explained further below).

{\bf Large injection:}
When the injected signal is much larger than the measured power spectrum, the data-only components can 
be neglected as many orders of magnitude smaller. We include a description of this regime for completeness in our discussion, but note that the upper limits that we compute are typically not determined by simulations in this regime (i.e. in using an empirical weighting scheme we've assumed the data to be dominated by foregrounds rather than the cosmological signal).  However, it is useful as a check of our system in a relatively simple case. As we can see from Figure \ref{fig:sigloss_terms}, the cross-terms (red) are small in comparison to the signal-only term (green). Here only does the signal-only term used in  \citetalias{ali_et_al2015} dominate the total power output. We again see that, in the estimated inverse covariance weighted case, the cross-terms behave as noise (positive and negative fluctuations around zero mean). This is for the same reason as at small injections --- here $\widehat{\textbf{C}}_{r}$ is dominated by the signal $\textbf{e}$. The cross-correlation can again be modeled as a symmetric distribution of zero mean and width $\sqrt{P_e}\sqrt{P_x}$.

@@

\begin{figure*}
	\centering
	\includegraphics[width=1\textwidth]{plots/sigloss_terms.pdf}
	\caption{Illustration of the power spectrum amplitude of various power spectrum terms as a function of injected EoR power level summed into the 
data. The net output ($P_{\rm out}=\widehat{\textbf{P}}_r$) is shown in black, the individual terms, as defined by Equation 
\eqref{eq:crossterm},  are shown in blue, red, green; the dotted black line indicates 1:1 input to output mapping.  The 
empirically estimated inverse covariance weighted case used in \citetalias{ali_et_al2015} is on the left and the uniform-weighted 
case is on the right. The details of the simulation used to generate the figure is explained in Section 
\ref{sec:Practice}; here we fit smooth polynomials to our data points to make an illustrative example.   The gray 
horizontal line is the power spectrum value of data alone (doesn't depend on injected power). The green signal-signal 
component is the term used in \citetalias{ali_et_al2015} to estimate signal loss. It is significantly higher 
than the net output because the cross-terms (red) are large and negative (black $=$ green $+$ red $+$ blue) . In the 
regime where cross-correlations between injection and data are not dominant, the cross-terms have a noise-like 
term with width $\sqrt{P_e}\sqrt{P_x}$. However, at power levels comparable to the data (the middle region), the cross-terms can produce large, negative signal due to the couplings between $\textbf{x}$ and $\textbf{e}$ which affect 
$\widehat{\textbf{C}}_{r}$. This causes the difference between the green curve (which exhibits negligible loss at 
the data level) and the black curve (which exhibits $\sim$$3$ orders of magnitude of loss at the data level for 
the weighted case). }
	\label{fig:sigloss_terms}
\end{figure*}

@@

Going forward, we treat every $k$-value separately in order to determine an upper limit on the EoR signal per $k$. We bin our simulation outputs along the $P_{\rm in}$ axis, and smooth the distribution of $P_{\rm out}$ values using kernel density estimators (\citealt{scott2008kernel}) for each bin (and normalizing them). Stitching all of them together results in a 2-dimensional smoothed transfer function, shown in Figure \ref{fig:sigloss_transfercurve} for both the weighted (left) and unweighted (right) cases. Although we only show figures for one $k$-value, we note that 
the shape of the transfer curve is similar for all $k$'s. This transfer function is the likelihood function in Bayes' rule, namely $\mathcal{L} (  P_{\rm out} | P_{\rm in})$.  We then invoke Bayes' interpretation and re-interpret it as the posterior $p(P_{\rm in}|P_{\rm out})$ where we recall that $P_{\rm out}$ is a model of our data. To do this we make a horizontal cut across at the data value $\widehat{\textbf{P}}_{x}$ (setting $P_{\rm out} = \widehat{\textbf{P}}_{x}$) to yield a posterior distribution for the signal. We compute the $95\%$ confidence interval (an upper limit on EoR) from these distributions.

@@

By-eye inspection of the transfer function in Figure \ref{fig:sigloss_transfercurve} gives a sense of what the signal loss result should be. The power spectrum value of our data, $
\widehat{\textbf{P}}_{x}$ is marked by the solid gray horizontal lines. From the left plot (empirically estimated inverse covariance weighting), one can eyeball that a data value of $10^{4} \,$mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, for example, would map approximately to an upper limit of $\sim10^{8} \,$mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$, implying a signal loss factor of $\sim10^{4}$. For the uniform-weighted case (right plot), we see no loss at a data value of $\sim10^{7} \,$mK$^{2}$ ($h^{-1}$ Mpc)$^{3}$.

@@

\begin{equation}
\label{eq:sense}
N(k) = \frac{X^{2}Y \Omega_{\rm eff} T_{\rm sys}^{2}}{\sqrt{2N_{\rm lst}N_{\rm seps}}\,t_{\rm int}N_{\rm days}N_{\rm bls}N_{\rm pols}}.
\end{equation}
We will now explain each factor in Equation \eqref{eq:sense} and highlight key differences from the numbers used in \citetalias{ali_et_al2015}.
